---
title: "Comparison of Existing and Optimized Edmonton Metro Shifts"
date: "February 15, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=12, fig.height=7)
```


```{r init, include=FALSE}




```


# Shift Datasets

This analysis compares the following datasets:

1. Historical: EDMO Metro Units deployed between Nov 1 2017 and Oct 31 2018. This set does not include IFT units assigned to 
   emergency events, or neighbouring community units assigned to Edmonton events.
1. Funded: EDMO Metro Shifts listed in the Jan 25, 2019 Provincial Profile.
1. Optimized Block Shifts (Scenario 1): An optimized weekly shift schedule was created using the 4 existing shift types observed in the Profile for EDMO Metro: i) 7-day, 12 hr Day/Night offset pairs ii) 7-day, 12 hr iii) 4-day, 10.5 hr and iv) 2-day, 10.5 hr. These shift types were fit to estimated weekly demand based on a FFT regression model (prediction interval=0.991), to minimize the total number of scheduled hours. Note: in this Block shift model, shifts are required to start the same time each day for consecutive days, matching current practice in the Profile. 
1. Optimized Free Shifts (Scenario 2): An optimized weekly shift schedule was created using the shift lengths 12, 10.5 and 8.4 hrs. Daily start times are independent of other days (no requirement for same start time across consecutive days). These shift types were fit to estimated weekly demand based on a FFT regression model (prediction interval=0.991), to minimize the total number of scheduled hours.


  
# Performance
  
## General