---
title: "Comparison of Existing and Optimized Edmonton Metro Shifts"
date: "February 15, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=12, fig.height=7)
```


```{r init, include=FALSE}

# Datasets
demand = drake::readd(demand)
histU = drake::readd(histU)
fundU = drake::readd(fundU)
crews.bl = drake::readd(crewsRmBl)
crews.fr = drake::readd(crewsRmFr)
crews1 = drake::readd(crewsExBl)
crews2 = drake::readd(crewsExFr)

# Params
duration.in.min = 15
myorigin = "2018-12-9 00:00:00" # A random sunday used to anchor the datetime column


# Convert
optBU = shifts$convert_shift_matrix(crews.bl$shift.matrix, myorigin, duration.in.min)
optFU = shifts$convert_shift_matrix(crews.fr$shift.matrix, myorigin, duration.in.min)
loptBU = shifts$convert_shift_matrix(crews1$shift.matrix, myorigin, duration.in.min)
loptFU = shifts$convert_shift_matrix(crews2$shift.matrix, myorigin, duration.in.min)

start.ts = min(demand$window)
end.ts = max(demand$window)
setkey(demand, window)

# Merge data with demand values
prep <- function(modU, weekly) {
  
  setkey(modU, window)

  # Stretch weekly values to year
  if(weekly) {
    modU = shifts$fill(modU, start.ts, end.ts)
  }
  
  # Join to demand
  modDT = shifts$shift_join(modU, demand)
  
  # Remove first/last day
  modDT = shifts$trim(modDT)
  
  # Add some utility columns
  modDT = shifts$weekly_bin(modDT, duration.in.min)
  
  return(modDT)
}

histU = prep(histU, FALSE)
fundU = prep(fundU, TRUE)
optBU = prep(optBU, TRUE)
optFU = prep(optFU, TRUE)
loptBU = prep(loptBU, TRUE)
loptFU = prep(loptFU, TRUE)

# Compute annual performance
histR = shifts$performance(histU)
fundR = shifts$performance(fundU)
optBR = shifts$performance(optBU)
optFR = shifts$performance(optFU)
loptBR = shifts$performance(loptBU)
loptFR = shifts$performance(loptFU)
rlist = list(histR, fundR, optBR, optFR, loptBR, loptFR)

# Compute weekly averages/metrics
histW = shifts$weekly_aggregate(histU, 60)
fundW = shifts$weekly_aggregate(fundU, 60)
optBW = shifts$weekly_aggregate(optBU, 60)
optFW = shifts$weekly_aggregate(optFU, 60)
loptBW = shifts$weekly_aggregate(loptBU, 60)
loptFW = shifts$weekly_aggregate(loptFU, 60)
wlist = list(histW, fundW, optBW, optFW, loptBW, loptFW)


# yearly counts
mlist = list(histU, fundU, optBU, optFU, loptBU, loptFU)

# Update start/end after trim operation
start.ts = min(histW$window)
end.ts = max(histW$window)


```


# Shift Datasets

This analysis compares the following datasets:

1. Historical: EDMO Metro Units deployed between Nov 1 2017 and Oct 31 2018. This set does not include IFT units assigned to 
   emergency events, or neighbouring community units assigned to Edmonton events.
1. Funded: EDMO Metro Shifts listed in the Jan 25, 2019 Provincial Profile.
1. Optimized Block Shifts (Scenario 1): An optimized weekly shift schedule was created using the 4 existing shift types observed in the Profile for EDMO Metro: i) 7-day, 12 hr Day/Night offset pairs ii) 7-day, 12 hr iii) 4-day, 10.5 hr and iv) 2-day, 10.5 hr. These shift types were fit to estimated weekly demand based on a FFT regression model (prediction interval=0.991), to minimize the total number of scheduled hours. Note: in this Block shift model, shifts are required to start the same time each day for consecutive days, matching current practice in the Profile. 
1. Optimized Free Shifts (Scenario 2): An optimized weekly shift schedule was created using the shift lengths 12, 10.5 and 8.4 hrs. Daily start times are independent of other days (no requirement for same start time across consecutive days). These shift types were fit to estimated weekly demand based on a FFT regression model (prediction interval=0.991), to minimize the total number of scheduled hours.
1. Optimized Block Shifts - Decreased Service Level: Same model as in Optimized Block Shifts, the service level was reduced so that total scheduled hours was similar to hours observed for historical.
1. Optimized Free Shifts - Decreased Service Level: Same model as in Optimized Free Shifts, the service level was reduced so that total scheduled hours was similar to hours observed for historical.

  
# Performance
  
## General

```{r perform1, warning=FALSE}

f1 <- function(modList, model.names) {
  thrs = sapply(modList, `[[`, 'StaffedTotalHrs')
  uhrs = sapply(modList, `[[`, 'UnitAvailableTotalHrs')
  upr = sapply(modList, `[[`, 'UnitAvailableProportion')
  ratio = uhrs / thrs
  
  df1 = data.frame(a=model.names, c=thrs, d=uhrs, e=upr, f=ratio)
  
  return(df1)
}

nms <- c('Historical', 'Funded', 'Optimized Block Shifts', 'Optimized Free Shifts', 'Optimized Block Shifts - Decreased Service Level', 'Optimized Free Shifts - Decreased Service Level')
df1 = f1(rlist, nms)

colnms <- c('Dataset','Total Scheduled Hours', 'Hours Unit Available', 'Proportion Unit Available', 'Hours Unit Available / Total Scheduled Hours' )
            
knitr::kable(df1, caption=paste("Performance Metrics for", start.ts,' - ', end.ts), row.names = FALSE, col.names = colnms) %>%
  kable_styling()

```

<style>
  .superbigimage{
    overflow-x:scroll;
    white-space: nowrap;
  }

.superbigimage img{
  max-width: none;
}
</style>

### Unit Availability Plots {.tabset}

#### Historical vs Optimized

Optimized Block Shifts using high service level to meet expected demand

<div class="superbigimage">
```{r plot1.01, warning=FALSE}

tmplist = list(histW, optBW)
nap = data.table(1-sapply(tmplist, `[[`, 'UnitAvailableProportion'))
thisnms = nms[c(1,3)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_bar(nap, thisnms, 
                                              title='Average Time Unit is Unavailable',
                                              ylabel = 'Proportion')
p

````
</div>

#### Historical vs Optimized 2

Optimized Block Shifts with reduced service level to match Historical staffed hours.

<div class="superbigimage">
```{r plot1.02, warning=FALSE}

tmplist = list(histW, loptBW)
nap = data.table(1-sapply(tmplist, `[[`, 'UnitAvailableProportion'))
thisnms = nms[c(1,5)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_bar(nap, thisnms, 
                                              title='Average Time Unit is Unavailable',
                                              ylabel = 'Proportion')
p

```
</div>

#### Historical vs Funded

Actual EDMO Metro Units deployed vs Units listed in Profile.

<div class="superbigimage">
```{r plot1.03, warning=FALSE}

tmplist = list(histW, fundW)
nap = data.table(1-sapply(tmplist, `[[`, 'UnitAvailableProportion'))
thisnms = nms[c(1,2)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_bar(nap, thisnms, 
                                              title='Average Time Unit is Unavailable',
                                              ylabel = 'Proportion')
p

```
</div>

#### Optimized Block Shifts vs Free Shifts

Free shifts are not constrainted by requiring the same consecutive multi-day start times.

<div class="superbigimage">
```{r plot1.04, warning=FALSE}

tmplist = list(optBW, optFW)
nap = data.table(1-sapply(tmplist, `[[`, 'UnitAvailableProportion'))
thisnms = nms[c(3,4)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_bar(nap, thisnms, 
                                              title='Average Time Unit is Unavailable',
                                              ylabel = 'Proportion')
p

```
</div>



### Unit Distribution Plots {.tabset}

#### Historical vs Optimized 

Optimized Block Shifts using high service level to meet expected demand

<div class="superbigimage">
```{r plot1.1, warning=FALSE}


weekly_count <- function(sDT) {
  cDT = sDT[,.(window=max(window), mean=mean(staffed)), .(weekly.bin, daily.bin)]
  cDT[,.(window, mean)]
}

histC = weekly_count(histU)
optBC = weekly_count(optBU)

counts = merge(histC, optBC, by='window', suffixes = c('.hist', '.opt'))
counts = shifts$weekly_bin(counts, duration.in.min)
setkeyv(counts, c("weekly.bin", "daily.bin"))

dt = seq(counts$window[1], counts$window[1]+lubridate::days(7), by=paste(duration.in.min, "mins"))
dt = dt[1:(length(dt)-1)]
counts[,window:=dt]
thisnms = nms[c(1,3)]
setnames(counts, c('mean.hist','mean.opt'), thisnms)

p = performance__plot_weekly_multi_metric_line(counts, thisnms,
                                               title='Average Staffed Units',
                                               ylabel = '# Units')
p

```
</div>

#### Historical vs Optimized 2

Optimized Block Shifts with reduced service level to match Historical staffed hours.

<div class="superbigimage">
```{r plot1.2, warning=FALSE}

loptBC = weekly_count(loptBU)

counts = merge(histC, loptBC, by='window', suffixes = c('.hist', '.opt'))
counts = shifts$weekly_bin(counts, duration.in.min)
setkeyv(counts, c("weekly.bin", "daily.bin"))
counts[,window:=dt]
thisnms = nms[c(1,5)]
setnames(counts, c('mean.hist','mean.opt'), thisnms)


p = performance__plot_weekly_multi_metric_line(counts, thisnms,
                                              title='Average Staffed Units',
                                              ylabel = '# Units')
p

```
</div>

#### Historical vs Funded

Actual EDMO Metro Units deployed vs Units listed in Profile.

<div class="superbigimage">
```{r plot1.3, warning=FALSE}

fundC = weekly_count(fundU)

counts = merge(histC, fundC, by='window', suffixes = c('.hist', '.opt'))
counts = shifts$weekly_bin(counts, duration.in.min)
setkeyv(counts, c("weekly.bin", "daily.bin"))
counts[,window:=dt]
thisnms = nms[c(1,2)]
setnames(counts, c('mean.hist','mean.opt'), thisnms)


p = performance__plot_weekly_multi_metric_line(counts, thisnms,
                                              title='Average Staffed Units',
                                              ylabel = '# Units')
p

```
</div>


#### Optimized Block Shifts vs Free Shifts

Free shifts are not constrainted by requiring the same consecutive multi-day start times.

<div class="superbigimage">
```{r plot1.4, warning=FALSE}

optFC = weekly_count(optFU)

counts = merge(optBC, optFC, by='window', suffixes = c('.hist', '.opt'))
counts = shifts$weekly_bin(counts, duration.in.min)
setkeyv(counts, c("weekly.bin", "daily.bin"))
counts[,window:=dt]
thisnms = nms[c(3,4)]
setnames(counts, c('mean.hist','mean.opt'), thisnms)


p = performance__plot_weekly_multi_metric_line(counts, thisnms,
                                              title='Average Staffed Units',
                                              ylabel = '# Units')
p

```
</div>

## Utilization

```{r perform2, warning=FALSE}

f2 <- function(modList, model.names, metric) {
  ave = sapply(modList, function(x) sapply(x[metric], `[[`, 'mean'))
  p50 = sapply(modList, function(x) sapply(x[metric], `[[`, 'p50'))
  p90 = sapply(modList, function(x) sapply(x[metric], `[[`, 'p90'))
  max = sapply(modList, function(x) sapply(x[metric], `[[`, 'max'))
  min = sapply(modList, function(x) sapply(x[metric], `[[`, 'min'))
  
  df = data.frame(a=model.names, b=ave, c=p50, d=p90, e=max, f=min)
  
  return(df)
}


df2 = f2(rlist, nms, 'Utilization')

colnms <- c("Dataset", "Mean", "Median", "90th Percentile", "Max", "Min")
knitr::kable(df2, caption=paste("Utilization Metrics for", start.ts,' - ', end.ts), row.names = FALSE, col.names = colnms) %>%
  kable_styling()


```


### Utilization Plots {.tabset}

#### Historical vs Optimized

Optimized Block Shifts using high service level to meet expected demand

<div class="superbigimage">
```{r plot2.01, warning=FALSE}

tmplist = list(histW, optBW)
nap = data.table(sapply(tmplist, `[[`, 'ut.mean'))
thisnms = nms[c(1,3)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_bar(nap, thisnms, 
                                              title='Average Utilization',
                                              ylabel = 'Proportion on Event')
p

````
</div>

#### Historical vs Optimized 2

Optimized Block Shifts with reduced service level to match Historical staffed hours.

<div class="superbigimage">
```{r plot2.02, warning=FALSE}

tmplist = list(histW, loptBW)
nap = data.table(sapply(tmplist, `[[`, 'ut.mean'))
thisnms = nms[c(1,5)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_bar(nap, thisnms, 
                                              title='Average Utilization',
                                              ylabel = 'Proportion on Event')
p

```
</div>

#### Historical vs Funded

Actual EDMO Metro Units deployed vs Units listed in Profile.

<div class="superbigimage">
```{r plot2.03, warning=FALSE}

tmplist = list(histW, fundW)
nap = data.table(sapply(tmplist, `[[`, 'ut.mean'))
thisnms = nms[c(1,2)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_bar(nap, thisnms, 
                                              title='Average Utilization',
                                              ylabel = 'Proportion on Event')
p

```
</div>

#### Optimized Block Shifts vs Free Shifts

Free shifts are not constrainted by requiring the same consecutive multi-day start times.

<div class="superbigimage">
```{r plot2.04, warning=FALSE}

tmplist = list(optBW, optFW)
nap = data.table(sapply(tmplist, `[[`, 'ut.mean'))
thisnms = nms[c(3,4)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_bar(nap, thisnms, 
                                              title='Average Utilization',
                                              ylabel = 'Proportion on Event')
p

```
</div>

## Availability

```{r perform3, warning=FALSE}

df3 = f2(rlist, nms, 'Availability')


knitr::kable(df3, caption=paste("Availability Metrics for", start.ts,' - ', end.ts), row.names = FALSE, col.names = colnms) %>%
  kable_styling()

```


Some of the largest difference between estimated and actual units occurred on:
  
  1. 2018-01-01 2018, Differences ranged from -10 to -2 over 00:30-1:30. New years.
  2. 2018-02-13, Differences ranged from -7 to -2 over 13:30-17:30. Freezing rain, lots of fall events (Card 17).
  3. 2018-10-11 Differences ranged from -6 to -3 over 16:30:00-18:15:00. No discovered explanation.
  

### Availability Plots {.tabset}

#### Historical vs Optimized

Optimized Block Shifts using high service level to meet expected demand

<div class="superbigimage">
```{r plot3.01, warning=FALSE}

tmplist = list(histW, optBW)
nap = data.table(sapply(tmplist, `[[`, 'av.mean'))
thisnms = nms[c(1,3)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_line(nap, thisnms, 
                                              title='Average Availability',
                                              ylabel = '# Units')
p

````
</div>

#### Historical vs Optimized 2

Optimized Block Shifts with reduced service level to match Historical staffed hours.

<div class="superbigimage">
```{r plot3.02, warning=FALSE}

tmplist = list(histW, loptBW)
nap = data.table(sapply(tmplist, `[[`, 'av.mean'))
thisnms = nms[c(1,5)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_line(nap, thisnms, 
                                              title='Average Availability',
                                              ylabel = '# Units')
p

```
</div>

#### Historical vs Funded

Actual EDMO Metro Units deployed vs Units listed in Profile.

<div class="superbigimage">
```{r plot3.03, warning=FALSE}

tmplist = list(histW, fundW)
nap = data.table(sapply(tmplist, `[[`, 'av.mean'))
thisnms = nms[c(1,2)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_line(nap, thisnms, 
                                              title='Average Availability',
                                              ylabel = '# Units')
p

```
</div>

#### Optimized Block Shifts vs Free Shifts

Free shifts are not constrainted by requiring the same consecutive multi-day start times.

<div class="superbigimage">
```{r plot3.04, warning=FALSE}

tmplist = list(optBW, optFW)
nap = data.table(sapply(tmplist, `[[`, 'av.mean'))
thisnms = nms[c(3,4)]
colnames(nap) = thisnms
nap = cbind(window=tmplist[[1]]$window, nap)

p = performance__plot_weekly_multi_metric_line(nap, thisnms, 
                                              title='Average Availability',
                                              ylabel = '# Units')
p

```
</div>


### Availability Distribution Plots {.tabset}

#### Historical vs Optimized

Optimized Block Shifts using high service level to meet expected demand

<div class="superbigimage">
```{r plot3.11, warning=FALSE}

i = c(1,3)
tmplist = mlist[i]
avDT = data.table(sapply(tmplist, `[[`, 'av'))
thisnms = nms[i]
colnames(avDT) = thisnms
avDT = cbind(window=tmplist[[1]]$window, avDT)

p = performance__plot_hour_histogram(avDT, thisnms, duration.in.min)

p

````
</div>

<div class="superbigimage">
```{r plot3.21, echo=FALSE}

p = performance__plot_availability_cumulative_distribution(avDT, thisnms)

p

```
</div>

#### Historical vs Optimized 2

Optimized Block Shifts with reduced service level to match Historical staffed hours.

<div class="superbigimage">
```{r plot3.12, warning=FALSE}

i = c(1,5)
tmplist = mlist[i]
avDT = data.table(sapply(tmplist, `[[`, 'av'))
thisnms = nms[i]
colnames(avDT) = thisnms
avDT = cbind(window=tmplist[[1]]$window, avDT)

p = performance__plot_hour_histogram(avDT, thisnms, duration.in.min)

p

```
</div>

<div class="superbigimage">
```{r plot3.22, echo=FALSE}

p = performance__plot_availability_cumulative_distribution(avDT, thisnms)

p

```
</div>

#### Historical vs Funded

Actual EDMO Metro Units deployed vs Units listed in Profile.

<div class="superbigimage">
```{r plot3.13, warning=FALSE}

i = c(1,2)
tmplist = mlist[i]
avDT = data.table(sapply(tmplist, `[[`, 'av'))
thisnms = nms[i]
colnames(avDT) = thisnms
avDT = cbind(window=tmplist[[1]]$window, avDT)

p = performance__plot_hour_histogram(avDT, thisnms, duration.in.min)

p

```
</div>

<div class="superbigimage">
```{r plot3.23, echo=FALSE}

p = performance__plot_availability_cumulative_distribution(avDT, thisnms)

p

```
</div>

#### Optimized Block Shifts vs Free Shifts

Free shifts are not constrainted by requiring the same consecutive multi-day start times.

<div class="superbigimage">
```{r plot3.14, warning=FALSE}

i = c(3,4)
tmplist = mlist[i]
avDT = data.table(sapply(tmplist, `[[`, 'av'))
thisnms = nms[i]
colnames(avDT) = thisnms
avDT = cbind(window=tmplist[[1]]$window, avDT)

p = performance__plot_hour_histogram(avDT, thisnms, duration.in.min)

p

```
</div>

<div class="superbigimage">
```{r plot3.24, echo=FALSE}

p = performance__plot_availability_cumulative_distribution(avDT, thisnms)

p

```
</div>
  
### Availability Distribution Table

```{r table3.3, echo=FALSE}

dts = lapply(mlist, function(m) { DT=t(performance__table_availability_cumulative_distribution(m)); colnames(DT)=DT[1,]; data.table(DT[-1,]) })
distnDT = rbindlist(dts, fill=TRUE)

keepcol = as.numeric(colnames(distnDT)) >= 0
distnDT = distnDT[,..keepcol]

distnDT = cbind(Metric=rep(c('Weekday', 'Weekend', '% Units', '% >= Units'), 6), Dataset=rep(nms, each=4), distnDT)

knitr::kable(distnDT, caption=paste("Unit Availability for", start.ts,' - ', end.ts), 
             row.names=FALSE, col.names=colnames(distnDT)) %>%
  kableExtra::kable_styling(bootstrap_options = "bordered") %>%
  kableExtra::column_spec(1:2, width = "15em; min-width: 15em;") %>%
  kableExtra::scroll_box(width = "800px", height = "100%")


```





  
