---
title: "comparemodels"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=12, fig.height=7)
```


```{r init, include=FALSE}

# Load datasets
demand = drake::readd(demand)
qmodU = drake::readd(qmodU)
rmodU = drake::readd(rmodU)

# Add datetime column
qmodU = queuemodel__convert(qmodU)
rmodU = queuemodel__convert(rmodU)

setkey(qmodU, window)
setkey(demand, window)
setkey(rmodU, window)

# Round regression targets
rmodU[,count:=ceiling(count)]

# Merge demand & unit data.tables
start.ts = min(demand$window)
end.ts = max(demand$window)

qmodU = shifts$fill(qmodU, start.ts, end.ts)
rmodU = shifts$fill(rmodU, start.ts, end.ts)

qmodDT = shifts$shift_join(qmodU, demand)
rmodDT = shifts$shift_join(rmodU, demand)

qmodDT = shifts$trim(qmodDT)
rmodDT = shifts$trim(rmodDT)

res.rmod = shifts$performance(rmodDT)
res.qmod = shifts$performance(qmodDT)

wk.rmod = shifts$weekly_aggregate(rmodDT, 60)
wk.qmod = shifts$weekly_aggregate(qmodDT, 60)

```


# Demand Modelling

## Models

Two methods were tried for fitting and estimating required units in each period given a service level (or prediction interval):

  1. Erlang-C queue model.
    - Inputs: Weekly Arrival rate and Service Time in each 15 minute period. 
    - Parameters: service level (probability of new event having to wait for unit to become available), 
        arrival rate lag (the number of preceeding periods to use in arrival rate average),
        service rate lag (the number of preceeding periods to use in arrival rate average),
        load threshold (when the unit estimate drops below this value, service rate is increased by 5% and a new unit 
                        requirement is calculated)
  2. Fourier Regression model.
    - Inputs: Number of concurrent units on event for each 15 minute period in the whole study duration
    - Parameters: prediction interval (the upper prediction interval bound determines the final required number of units)
  
## Performance
  
### General
  
  
```{r perform1, warning=FALSE}

rownms <- c('Queue Model','FFT Regression Model')

colnms <- c('Total Hours', "Hours Unit Available", "Proportion Unit Available")
df1 = data.frame(A=c(res.qmod$StaffedTotalHrs, res.rmod$StaffedTotalHrs), 
                 B=c(res.qmod$UnitAvailableTotalHrs, res.rmod$UnitAvailableTotalHrs),
                 C=c(res.qmod$UnitAvailableProportion, res.rmod$UnitAvailableProportion),
                 row.names = rownms)
start.ts = min(qmodDT$window)
end.ts = max(qmodDT$window)              
knitr::kable(df1, caption=paste("Performance Metrics for", start.ts,' - ', end.ts), row.names = TRUE, col.names = colnms)

```

<style>
  .superbigimage{
    overflow-x:scroll;
    white-space: nowrap;
  }

.superbigimage img{
  max-width: none;
}
</style>
  

<div class="superbigimage">
```{r plot1, warning=FALSE}

nap = wk.rmod[wk.qmod, .(window=window, fft.regression=1-UnitAvailableProportion, queue.model=1-i.UnitAvailableProportion)]

p = performance__plot_weekly_multi_metric_bar(nap, c('fft.regression','queue.model'), title='Average Time Unit is Unavailable',
                                               ylabel = 'Proportion')
p

````
</div>

### Utilization

```{r perform2, warning=FALSE}


colnms <- c("Mean", "Median", "90th Percentile", "Max", "Min")
df2 = data.frame(A=c(res.qmod$Utilization$mean, res.rmod$Utilization$mean),
                 B=c(res.qmod$Utilization$p50, res.rmod$Utilization$p50),
                 C=c(res.qmod$Utilization$p90, res.rmod$Utilization$p90),
                 D=c(res.qmod$Utilization$max, res.rmod$Utilization$max),
                 E=c(res.qmod$Utilization$min, res.rmod$Utilization$min),
                 row.names = rownms)
knitr::kable(df2, caption=paste("Utilization Metrics for", start.ts,' - ', end.ts), row.names = TRUE, col.names = colnms)

```
<div class="superbigimage">
```{r plot2} 

cols = c('fft.regression','queue.model')
ut = wk.rmod[wk.qmod, .(window=window, fft.regression=ut.mean, queue.model=i.ut.mean)]

p = performance__plot_weekly_multi_metric_line(ut, c('fft.regression','queue.model'), title='Average Utilization',
                                               ylabel = 'Proportion on Event')
p
```
</div>

### Availability

```{r perform3, warning=FALSE}


colnms <- c("Mean", "Median", "90th Percentile", "Max", "Min")
df2 = data.frame(A=c(res.qmod$Availability$mean, res.rmod$Availability$mean),
                 B=c(res.qmod$Availability$p50, res.rmod$Availability$p50),
                 C=c(res.qmod$Availability$p90, res.rmod$Availability$p90),
                 D=c(res.qmod$Availability$max, res.rmod$Availability$max),
                 E=c(res.qmod$Availability$min, res.rmod$Availability$min),
                 row.names = rownms)
knitr::kable(df2, caption=paste("Availability Metrics for", start.ts,' - ', end.ts), row.names = TRUE, col.names = colnms)

```


Some of the largest difference between estimated and actual units occurred on:
  
  1. 2018-01-01 2018, Differences ranged from -10 to -2 over 00:30-1:30. New years.
  2. 2018-02-13, Differences ranged from -7 to -2 over 13:30-17:30. Freezing rain, lots of fall events (Card 17).
  3. 2018-10-11 Differences ranged from -6 to -3 over 16:30:00-18:15:00. No discovered explanation.
  
  
<div class="superbigimage">
```{r plot3} 

cols = c('fft.regression','queue.model')
av = wk.rmod[wk.qmod, .(window=window, fft.regression=av.mean, queue.model=i.av.mean)]

p = performance__plot_weekly_multi_metric_line(av, c('fft.regression','queue.model'), title='Average Availability',
                                               ylabel = '# Units')
p
```
</div>
  

### Model Evaluation
  
```{r perform4, echo=FALSE, warning=FALSE}
  
  
  colnms <- c("Overcapacity Mean", "Overcapacity Standard Deviation", "Overcapacity RMSE", "Overcapacity Total",
              "Undercapacity Mean", "Undercapacity Standard Deviation", "Undercapacity RMSE", "Undercapacity Total")
  df3 = data.frame(A=c(res.qmod$OvercapacityUnitHrs$mean, res.rmod$OvercapacityUnitHrs$mean),
                   B=c(res.qmod$OvercapacityUnitHrs$sd, res.rmod$OvercapacityUnitHrs$sd),
                   C=c(res.qmod$OvercapacityUnitHrs$rmse, res.rmod$OvercapacityUnitHrs$rmse),
                   D=c(res.qmod$OvercapacityUnitHrs$tot, res.rmod$OvercapacityUnitHrs$tot),
                   E=c(res.qmod$UndercapacityUnitHrs$mean, res.rmod$UndercapacityUnitHrs$mean),
                   F=c(res.qmod$UndercapacityUnitHrs$sd, res.rmod$UndercapacityUnitHrs$sd),
                   G=c(res.qmod$UndercapacityUnitHrs$rmse, res.rmod$UndercapacityUnitHrs$rmse),
                   H=c(res.qmod$UndercapacityUnitHrs$tot, res.rmod$UndercapacityUnitHrs$tot),
                   row.names = rownms)
  knitr::kable(df3, caption=paste("Error Metrics for", start.ts,' - ', end.ts), row.names = TRUE, col.names = colnms)
  
```

To assess the models, error was split into Overcapacity (Unit hours below demand level i.e. Code Red) and Undercapacity (Unit hours in excess of demand).
