---
title: "Comparison of Queue and Regression-based Modelling"
date: "February 13, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=12, fig.height=7)
```


```{r init, include=FALSE}

# Load datasets
demand = drake::readd(demand)
# Load data into memory
ar2 = drake::readd(ar2)
st2 = drake::readd(st2)
modQ = modeltest__makeQ(ar2, st2)
modR = modeltest__makeR(demand)


start.ts = min(demand$window)
end.ts = max(demand$window)
setkey(demand, window)

# Round regression targets
modR = lapply(modR, function(rmodU) rmodU[,s:=ceiling(s)])

# Merge data with demand values
prep <- function(modU) {
  # Add datetime column
  modU = queuemodel__convert(modU)
  setkey(modU, window)

  # Stretch weekly values to year
  modU = shifts$fill(modU, start.ts, end.ts)
  
  # Join to demand
  modDT = shifts$shift_join(modU, demand)
  
  # Remove first/last day
  modDT = shifts$trim(modDT)
  
  return(modDT)
}

modQ = lapply(modQ, prep)
modR = lapply(modR, prep)

# Compute performance
res.modQ = lapply(modQ, shifts$performance)
res.modR = lapply(modR, shifts$performance)

# Compute weekly averages/metrics for best case
wk.rmod = shifts$weekly_aggregate(modR[['p0.991']], 60)
wk.qmod = shifts$weekly_aggregate(modQ[['p0.001']], 60)

# Yearly stats
rmod = modR[['p0.991']]
qmod = modQ[['p0.001']]

start.ts = min(rmod$window)
end.ts = max(rmod$window)


```


# Models

Two methods were tried for fitting and estimating required units in each period given a service level (or prediction interval) in a single week (our planning duration):

  1. Erlang-C queue model.
    - Inputs: Weekly Arrival rate and Service Time in each 15 minute period. 
    - Parameters: service level (probability of new event having to wait for unit to become available), 
        arrival rate lag (the number of preceeding periods to use in arrival rate average),
        service rate lag (the number of preceeding periods to use in arrival rate average),
        load threshold (when the unit estimate drops below this value, service rate is increased by 5% and a new unit 
                        requirement is calculated)
  2. Fourier Regression model.
    - Inputs: Number of concurrent units on event for each 15 minute period in the whole study duration
    - Parameters: prediction interval (the upper prediction interval bound determines the final required number of units)
  
# Performance
  
## General
  
  
```{r perform1, warning=FALSE}

f1 <- function(res.modQ, res.modR) {
  model = c(rep('Queue Model', length(res.modQ)), rep('FFT Regression Model', length(res.modR)))
  sl = c(names(res.modQ), names(res.modR))
  thrs = c(sapply(res.modQ[names(res.modQ)], `[[`, 'StaffedTotalHrs'),
           sapply(res.modR[names(res.modR)], `[[`, 'StaffedTotalHrs'))
  uhrs = c(sapply(res.modQ[names(res.modQ)], `[[`, 'UnitAvailableTotalHrs'),
           sapply(res.modR[names(res.modR)], `[[`, 'UnitAvailableTotalHrs'))
  upr = c(sapply(res.modQ[names(res.modQ)], `[[`, 'UnitAvailableProportion'),
           sapply(res.modR[names(res.modR)], `[[`, 'UnitAvailableProportion'))
  
  df1 = data.frame(a=model, b=sl, c=thrs, d=uhrs, f=upr)
  
  return(df1)
}

df1 = f1(res.modQ, res.modR)

colnms <- c('Model','Service Level', 'Total Hours', 'Hours Unit Available', 'Proportion Unit Available')
            
knitr::kable(df1, caption=paste("Performance Metrics for", start.ts,' - ', end.ts), row.names = FALSE, col.names = colnms) %>%
  kable_styling()

```

<style>
  .superbigimage{
    overflow-x:scroll;
    white-space: nowrap;
  }

.superbigimage img{
  max-width: none;
}
</style>
  

<div class="superbigimage">
```{r plot1, warning=FALSE, fig.cap='*Showing service level p0.001 in queue model and p0.991 prediction interval for regression model'}

nap = wk.rmod[wk.qmod, .(window=window, fft.regression=1-UnitAvailableProportion, queue.model=1-i.UnitAvailableProportion)]

p = performance__plot_weekly_multi_metric_bar(nap, c('fft.regression','queue.model'), title='Average Time Unit is Unavailable',
                                               ylabel = 'Proportion')
p

````
</div>

  
<div class="superbigimage">
```{r plot1.1, warning=FALSE, fig.width=12, fig.height=12}

ggplot(data=df1, aes(x=c, y=f, group=a)) +
  geom_line(aes(color=a,linetype=a)) +
  geom_point(aes(shape=a)) +
  geom_text(aes(label=b),hjust=1, vjust=1) +
  labs(y="Proportion Unit Available", x="Annual Unit Hours") + 
  theme(panel.background = element_rect(fill="#ffffff", colour=NA),
        axis.title.x = element_text(vjust=-0.2), axis.title.y=element_text(vjust=1.5),
        title = element_text(vjust=1.2),
        panel.border = element_blank(),
        axis.line =element_line(size=0.5, colour="grey80"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(size=0.5, colour="grey80"),
        panel.grid.major.y = element_line(size=0.5, colour="grey80"),
        axis.ticks = element_blank(),
        legend.position = "bottom", 
        axis.title = element_text(size=rel(0.8)),
        strip.text = element_text(size=rel(1)),
        strip.background=element_rect(fill="#ffffff", colour=NA),
        panel.spacing.y = unit(1.5, "lines"),
        legend.key = element_blank(),
        legend.title=element_blank())


````
</div>
  

## Utilization

```{r perform2, warning=FALSE}

f2 <- function(res.modQ, res.modR, metric) {
  model = c(rep('Queue Model', length(res.modQ)), rep('FFT Regression Model', length(res.modR)))
  sl = c(names(res.modQ), names(res.modR))
  ave = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'mean')),
          sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'mean')))
  p50 = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'p50')),
          sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'p50')))
  p90 = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'p90')),
          sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'p90')))
  max = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'max')),
          sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'max')))
  min = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'min')),
          sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'min')))
  
  df = data.frame(a=model, b=sl, c=ave, d=p50, e=p90, f=max, g=min)
  
  return(df)
}

df2 = f2(res.modQ, res.modR, 'Utilization')

colnms <- c("Model", "Service Level", "Mean", "Median", "90th Percentile", "Max", "Min")
knitr::kable(df2, caption=paste("Utilization Metrics for", start.ts,' - ', end.ts), row.names = FALSE, col.names = colnms) %>%
  kable_styling()


```

<div class="superbigimage">
```{r plot2, fig.cap='*Showing service level p0.001 in queue model and p0.991 prediction interval for regression model'} 

cols = c('fft.regression','queue.model')
ut = wk.rmod[wk.qmod, .(window=window, fft.regression=ut.mean, queue.model=i.ut.mean)]

p = performance__plot_weekly_multi_metric_line(ut, c('fft.regression','queue.model'), title='Average Utilization',
                                               ylabel = 'Proportion on Event')
p
```
</div>
  

## Availability

```{r perform3, warning=FALSE}

df3 = f2(res.modQ, res.modR, 'Availability')

colnms <- c("Model", "Service Level", "Mean", "Median", "90th Percentile", "Max", "Min")

knitr::kable(df3, caption=paste("Availability Metrics for", start.ts,' - ', end.ts), row.names = FALSE, col.names = colnms) %>%
  kable_styling()

```


Some of the largest difference between estimated and actual units occurred on:
  
  1. 2018-01-01 2018, Differences ranged from -10 to -2 over 00:30-1:30. New years.
  2. 2018-02-13, Differences ranged from -7 to -2 over 13:30-17:30. Freezing rain, lots of fall events (Card 17).
  3. 2018-10-11 Differences ranged from -6 to -3 over 16:30:00-18:15:00. No discovered explanation.
  
  
<div class="superbigimage">
```{r plot3, fig.cap='*Showing service level p0.001 in queue model and p0.991 prediction interval for regression model'} 

cols = c('fft.regression','queue.model')
av = wk.rmod[wk.qmod, .(window=window, fft.regression=av.mean, queue.model=i.av.mean)]

p = performance__plot_weekly_multi_metric_line(av, cols, title='Average Availability',
                                               ylabel = '# Units')
p
```
</div>
  

<div class="superbigimage">
```{r plot3.1, echo=FALSE}

duration.in.min = 15

avDT = rmod[qmod, .(window=window, fft.regression=av, queue.model=i.av)]
cols = c('fft.regression','queue.model')

p = performance__plot_hour_histogram(avDT, cols, duration.in.min)

p

```
</div>
  

<div class="superbigimage">
```{r plot3.2, echo=FALSE}

p = performance__plot_availability_cumulative_distribution(avDT, cols)

p

```
</div>
  


```{r table3.3, echo=FALSE}

dt1 = performance__table_availability_cumulative_distribution(qmod)
dt2 = performance__table_availability_cumulative_distribution(rmod)

setkey(dt1, av)
setkey(dt2, av)

avdt = merge(dt1, dt2, all=TRUE)
avdt = avdt[av >= 0]

avdt = cbind(c('', rep(c('Weekday', 'Weekend', '% Units', '% >= Units'),2)),
             c('', rep('Queue Model',4), rep('FFT Regression Model',4)),
             t(avdt))

ord = c(4,8,5,9,2,6,3,7)
avdtbody = avdt[ord,]

nc = ncol(avdtbody)

knitr::kable(as.data.frame(avdtbody), caption=paste("Unit Availability for", start.ts,' - ', end.ts), row.names=FALSE, col.names=avdt[1,]) %>%
  kableExtra::kable_styling(bootstrap_options = "bordered") %>%
  kableExtra::column_spec(1:2, width = "12em; min-width: 12em;") %>%
  kableExtra::column_spec(3:nc, width = "2.5em; min-width: 2.5em;") %>%
  kableExtra::scroll_box(width = "800px", height = "100%")
  

```

  

## Model Evaluation
  
```{r perform4, echo=FALSE, warning=FALSE}

  f3 <- function(res.modQ, res.modR, metric) {
    model = c(rep('Queue Model', length(res.modQ)), rep('FFT Regression Model', length(res.modR)))
    sl = c(names(res.modQ), names(res.modR))
    ave = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'mean')),
            sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'mean')))
    sd = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'sd')),
            sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'sd')))
    rmse = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'rmse')),
            sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'rmse')))
    tot = c(sapply(res.modQ[names(res.modQ)], function(x) sapply(x[metric], `[[`, 'tot')),
             sapply(res.modR[names(res.modR)], function(x) sapply(x[metric], `[[`, 'tot')))
    
    df = data.frame(a=model, b=sl, c=ave, d=sd, e=rmse, f=tot)
    
    return(df)
  }

  oc = f3(res.modQ, res.modR, 'OvercapacityUnitHrs')
  uc = f3(res.modQ, res.modR, 'UndercapacityUnitHrs')
  df4= cbind(oc,uc[c('c','d','e', 'f')])
  
  colnms <- c("Model", "Service Level", "Overcapacity Mean", "Overcapacity Standard Deviation", "Overcapacity RMSE", "Overcapacity Total",
              "Undercapacity Mean", "Undercapacity Standard Deviation", "Undercapacity RMSE", "Undercapacity Total")

  knitr::kable(df4, caption=paste("Error Metrics for", start.ts,' - ', end.ts), row.names = FALSE, col.names = colnms) %>%
    kable_styling()
  
```

To assess the models, error was split into Overcapacity (Unit hours below demand level i.e. Code Red) and Undercapacity (Unit hours in excess of demand).
